---
title: "Workshop.4"
author: "Sydney Russo"
date: "2025-03-19"
output: html_document
---

Workshop 1

uploading libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(hexbin)
#install.packages('viridis')
```


Creating the first plot
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour=class))
#this graph shows a negative relationship between engine size and fuel efficiency

#changing the appearance of the graph
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

#what happens if you map an aesthetic to something other than a variable name
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

Using Facet Wraps
```{r}
ggplot(data=mpg) + 
  geom_point(mapping =aes(x = displ, y = hwy)) +
  facet_wrap (~class, nrow = 2)
```

Using Facet Wrap Grids
```{r}
#this allows the facet to split by more than one variable

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, colour = class))+
  facet_grid(drv ~ cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, colour = class))+
  facet_grid(.~ cyl)
```
Exercise
#what does nrow do? what does ncol do? what other options control the layout of the individual panels?
#nrow and ncol are the number or rows and columns that are present
#there are scales, shrinking (will shrink scales to fit output of statistics not raw data if true, if false it will be a range of raw data before statistical summary), switch function allows for the X and Y to switch to top, bottom, right or left. there are many more aswell.


Fitting lines
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes (x=displ, y=hwy))

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy))+
  geom_smooth(mapping = aes(x=displ, y=hwy, linetype = drv))
#setting the group aesthetic
ggplot(data = mpg)+
  geom_smooth(mapping=aes(x=displ, y=hwy, group=drv))

#changing the colour
ggplot(data=mpg)+
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend=FALSE)

#mapping multiple geoms on one plot
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  geom_smooth(mapping = aes(x = displ, y = hwy))

#an easier way to do above if I wanted to change anything
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point()+
  geom_smooth()

#adding colors
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth(data = filter (mpg, class =="subcompact"), se =FALSE)
```

Exercise
#1. line: geom_line  , Boxplot:geom_boxplot  , Histogram: geom_histogram, an area chart: geom_area
#2. I predict that the graphs will all look different becuase they are asking for the data to be laid out in different forms.
```{r warning=FALSE}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_line()
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_boxplot()
```

#3. I would assume that these two graphs are saying the same thing just in 2 different ways. and yes they are the same.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point()+
  geom_smooth()
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  geom_smooth(mapping = aes(x = displ, y = hwy))
```


Plotting Statistics
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x =cut))
```

Overriding defaults in ggplot
```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
demo

diamonds
ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")

diamonds
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, y = stat(prop), group = 1))

```

Plotting Statistical details
```{r}
ggplot(data=diamonds)+
  stat_summary(
    mapping =aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )

#making it pretty
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds)+
  geom_bar(mapping = aes (x = cut, fill = cut))
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill = clarity))

#altering transparency
ggplot(data=diamonds, mapping = aes(x=cut, fill=clarity))+
  geom_bar(alpha=1/5, position = "identity")

#to color the bar outlines with no fill color
ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "identity")

ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "fill")


ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "dodge")


ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "jitter")

```



WORKSHOP 2

Labels
```{r warning=FALSE}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(title = "Fuel efficiency generally decreases with engine size")
#subtitle adds additional detail in a smaller font beneath the title and caption adds text at the bottom right of the plot
#caption adds text at the bottom right of the plot, often used to describe the source of the data

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(title = "Fuel efficiency generally decreases with engine size",
       subtitle = "Two seaters (sports cars) are an exception becuse of their light weight",
       caption = "Data from Feuleconomy.gov")

#you can also use labs to change the labels on the axis and legend titles         
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(
    x = "Engine displacement (L)",
    y = "Highway fueleconomy (mpg)",
    colour = "car type"
  )

```

Annotations
```{r}
best_in_class <- mpg |>
  group_by(class) |>
  filter(row_number(desc(hwy))==1)

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_text(aes(label = model), data = best_in_class)
```

Scales
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color = class))+
  scale_x_continuous()+
  scale_y_continuous()+
  scale_color_discrete()
```

Axis Ticks
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point()+
  scale_y_continuous(breaks = seq(15,40, by = 5))

ggplot(mpg, aes(displ, hwy))+
  geom_point()+
  scale_x_continuous(labels = NULL)+
  scale_y_continuous(labels = NULL)
```

Legends and color schemes
```{r}
base <- ggplot (mpg, aes(displ, hwy))+
  geom_point(aes(color=class))
base + theme(legend.position ="left")
base+theme(legend.position = "top")
base+theme(legend.position = "bottom")
base+theme(legend.position = "right") #this is the default legend position
base+theme(legend.position = "NONE")
```

Replacing a scale
```{r}
ggplot (diamonds, aes(carat, price)) +
  geom_bin2d()+
  scale_x_log10()+
  scale_y_log10()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=drv))+
  scale_color_brewer(palette = "set1")

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=drv, shape = drv))+
  scale_color_brewer(palette = "set1")

presidential|>
  mutate(id =33 + row_number()) |>
  ggplot(aes(start, id, color=party))+
    geom_point()+
    geom_segment(aes(xend = end, yend = id))+
    scale_color_manual(values=c(Republican = "red", Democratic = "blue"))
  
```
```{r}
df <- tibble( #not we are just making a fake dataset so we can plot it
  x= rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() + 
  coord_fixed()
ggplot(df, aes(x,y))+
  geom_hex()+
  viridis::scale_fill_viridis()+
  coord_fixed()


```
Themes
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_bw()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_light()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_classic()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_dark()

theme(panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size=8),
      panel.grid.major = element_blank(),
      axis.text.y=element_text(color="black"),
      axis.text.x=element_text(color="black"),
      text=element_text(family="Arial"))

```

Saving and Exporting your Plots
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()

ggsave("my-plot.pdf")
```
Workshop 2 Assignment
```{r}
Qfish <-read.csv("../data/QFish_Data.csv")
print(Qfish)

```
Plotting the data, it is mainly cleaned, there are no NAs or missing values
```{r}
library(ggplot2)
ggplot(Qfish, aes (x = Area, y = X2001.Total))+
  geom_bar(stat = "identity", fill = "darkcyan") +
  theme_minimal()+
  labs(title = "Total Catch in 2001 by Area",
       x = NULL,
       y = "Total Cathc") +
  theme(axis.text.x =element_text(angle = 45, hjust = 1, size = 6))


```


To Import the dataset: My first step for importing the dataset was to export a .csv file from Qfish. Once I have it in my correct file I then used the read.csv and the path to import the file into Rstudio.To make what I wanted easier to work with I created a new .csv file with only total 2001 data and the area.
Cleaning the data: I fixed the columns with the proper headers and created a new dataset with the data I needed
Creating my plot: I decided to make a bar graph as that was the best way to show the total number of species and animals caught in the year 2001. I created the boxplot and played around with color, phone sizes and axis labels. 

Workshop 3

```{r}
library(tidyverse)
table1
library(ggplot2)
ggplot(table1, aes(year, cases))+
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))
```

```{r}
table1

table2
table3
```
1. Table 1 is set up well with country, year cases and population. This shows the population per year per country with the number of cases. table2 shows the count of cases and population in each country separated by year. table3 has the rate in there but as the growth rate per year in each country but it has not been calculated yet.
2. a. using table 1 we would use a combination of the pipe function and the pivot longer to call the cases with country and year because there are 2 years for each country. and use the mutate function to put it into a new row in the table
b. and then to do the matching population per country per year, we would do the same as above but instead of case it would be population and then would mutate it so that it created a new row in the chart
c. using those new columns and creating a new column I can then calculate the the case by population and multiply it by 10,000 and use the mutate function to put that back into the dataset.

```{r}
billboard |>
  pivot_longer(
    cols = starts_with("wk"),
    names_to = "week",
    values_to = "rank",
    values_drop_na = TRUE
  )

```

```{r}
df <- tribble (
  ~id, ~bp1, ~bp2,
  "A", 100, 120,
  "B", 140, 115,
  "C", 120, 125
)
#this is creating out own dataset called df with 3 variables

df|>
  pivot_longer(
    cols = bp1:bp2,
    names_to = "measurement",
    values_to ="value"
  )
#the above took the values for both bp1 and bp2 into one column and then another column was its matched bp1 or bp2
```

```{r}
cms_patient_experience |>
  pivot_wider(
    id_cols = starts_with("org"),
    names_from = measure_cd,
    values_from = prf_rate
  )
```
```{r}
df <- tribble (
  ~id, ~measurement, ~value,
  "A", "bp1", 100,
  "B", "bp2", 115,
  "B", "bp1", 140,
  "A", "bp2", 120,
  "A", "bp3", 105
)
print(df)

df |>
  pivot_wider(
    names_from = measurement,
    values_from = value
  )

df |>
  distinct (measurement) |>
  pull()

df |>
  select(-measurement, -value) |>
  distinct()|>
  mutate(x = NA, y = NA, z = NA)


```

Excercise 4.5.5
```{r}
stocks <- tibble(
  year = c(2015, 2015, 2016, 2016),
  half = c(1, 2, 1, 2),
  return = c(1.88, 0.59, 0.92, 0.17)
)

stocks %>%
  pivot_wider(names_from = year, values_from = return) %>%
  pivot_longer('2015':'2016',names_to = "year", values_to = "return")
```

1. it is becuase they sort the data differently. One organizes by 1/2 1 na dthe other by 1/2 2

2. the code fails because the years need the quotations around them

3. this tibble would need to be longer. the variables are pregnant, male and female


```{r}
table3 |>
  separate(rate, into = c("cases", "population"), convert = TRUE)

table3 |>
  separate(year, into = c("century", "year"), sep = 2)


table5|>
  unite(new,century,year,sep = "")

```

```{r}
treatment <- tribble(
  ~person, ~treatment, ~response,
  "Derrick", 1, 7, 
  NA, 2, 10,
  NA, 3, NA,
  "Katherine", 1, 4
)

#you can use fill in these missing values with tidyr::fill(), it works like select()
treatment |>
  fill(everything()) 
#this treatment is sometimes calles "last observation carried forward"

```

```{r}
x <- c(1,4,5,7,NA)
coalesce(x,0)

x <- c(1,4,5,7,-99)
na_if(x,-99)

x<- c (NA, NaN)
x*10

x==1

is.na(x)
```

```{r}
stocks <- tibble(
  year = c(2020,2020,2020,2020,2021,2021,2021),
  qrt = c(1,2,3,4,2,3,4),
  price = c(1.88,0.59,0.35,NA,0.92,0.17,2.66)
)

stocks |>
  pivot_wider(
    names_from = qrt,
    values_from = price
  )
```
PAGE &% THE BOTTOM





Exercise 4.7.3


1. each of these are creating a file.
  1. there are not enough column names
  2. there are too many rows for the amount of columns present
  3. there is a quotation issue with this one
  4. there is an issue with number and character 
  5. the last one needs a comma not a semicolon











