---
title: "Workshop.4"
author: "Sydney Russo"
date: "2025-03-19"
output: html_document
---

Workshop 1

uploading libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(hexbin)
#install.packages('viridis')
```


Creating the first plot
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour=class))
#this graph shows a negative relationship between engine size and fuel efficiency

#changing the appearance of the graph
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

#what happens if you map an aesthetic to something other than a variable name
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

Using Facet Wraps
```{r}
ggplot(data=mpg) + 
  geom_point(mapping =aes(x = displ, y = hwy)) +
  facet_wrap (~class, nrow = 2)
```

Using Facet Wrap Grids
```{r}
#this allows the facet to split by more than one variable

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, colour = class))+
  facet_grid(drv ~ cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, colour = class))+
  facet_grid(.~ cyl)
```

Exercise
#what does nrow do? what does ncol do? what other options control the layout of the individual panels?
#nrow and ncol are the number or rows and columns that are present
#there are scales, shrinking (will shrink scales to fit output of statistics not raw data if true, if false it will be a range of raw data before statistical summary), switch function allows for the X and Y to switch to top, bottom, right or left. there are many more aswell.


Fitting lines
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes (x=displ, y=hwy))

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy))+
  geom_smooth(mapping = aes(x=displ, y=hwy, linetype = drv))
#setting the group aesthetic
ggplot(data = mpg)+
  geom_smooth(mapping=aes(x=displ, y=hwy, group=drv))

#changing the colour
ggplot(data=mpg)+
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend=FALSE)

#mapping multiple geoms on one plot
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  geom_smooth(mapping = aes(x = displ, y = hwy))

#an easier way to do above if I wanted to change anything
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point()+
  geom_smooth()

#adding colors
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth(data = filter (mpg, class =="subcompact"), se =FALSE)
```

Exercise
#1. line: geom_line  , Boxplot:geom_boxplot  , Histogram: geom_histogram, an area chart: geom_area
#2. I predict that the graphs will all look different becuase they are asking for the data to be laid out in different forms.
```{r warning=FALSE}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_line()
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_boxplot()
```

#3. I would assume that these two graphs are saying the same thing just in 2 different ways. and yes they are the same.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+
  geom_point()+
  geom_smooth()
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  geom_smooth(mapping = aes(x = displ, y = hwy))
```


Plotting Statistics
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x =cut))
```

Overriding defaults in ggplot
```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
demo

diamonds
ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")

diamonds
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, y = stat(prop), group = 1))

```

Plotting Statistical details
```{r}
ggplot(data=diamonds)+
  stat_summary(
    mapping =aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )

#making it pretty
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds)+
  geom_bar(mapping = aes (x = cut, fill = cut))
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill = clarity))

#altering transparency
ggplot(data=diamonds, mapping = aes(x=cut, fill=clarity))+
  geom_bar(alpha=1/5, position = "identity")

#to color the bar outlines with no fill color
ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "identity")

ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "fill")


ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "dodge")


ggplot(data=diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position = "jitter")

```



WORKSHOP 2

Labels
```{r warning=FALSE}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(title = "Fuel efficiency generally decreases with engine size")
#subtitle adds additional detail in a smaller font beneath the title and caption adds text at the bottom right of the plot
#caption adds text at the bottom right of the plot, often used to describe the source of the data

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(title = "Fuel efficiency generally decreases with engine size",
       subtitle = "Two seaters (sports cars) are an exception becuse of their light weight",
       caption = "Data from Feuleconomy.gov")

#you can also use labs to change the labels on the axis and legend titles         
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se.e = FALSE)+
  labs(
    x = "Engine displacement (L)",
    y = "Highway fueleconomy (mpg)",
    colour = "car type"
  )

```

Annotations
```{r}
best_in_class <- mpg |>
  group_by(class) |>
  filter(row_number(desc(hwy))==1)

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_text(aes(label = model), data = best_in_class)
```

Scales
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color = class))+
  scale_x_continuous()+
  scale_y_continuous()+
  scale_color_discrete()
```

Axis Ticks
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point()+
  scale_y_continuous(breaks = seq(15,40, by = 5))

ggplot(mpg, aes(displ, hwy))+
  geom_point()+
  scale_x_continuous(labels = NULL)+
  scale_y_continuous(labels = NULL)
```

Legends and color schemes
```{r}
base <- ggplot (mpg, aes(displ, hwy))+
  geom_point(aes(color=class))
base + theme(legend.position ="left")
base+theme(legend.position = "top")
base+theme(legend.position = "bottom")
base+theme(legend.position = "right") #this is the default legend position
base+theme(legend.position = "NONE")
```

Replacing a scale
```{r}
ggplot (diamonds, aes(carat, price)) +
  geom_bin2d()+
  scale_x_log10()+
  scale_y_log10()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=drv))+
  scale_color_brewer(palette = "set1")

ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=drv, shape = drv))+
  scale_color_brewer(palette = "set1")

presidential|>
  mutate(id =33 + row_number()) |>
  ggplot(aes(start, id, color=party))+
    geom_point()+
    geom_segment(aes(xend = end, yend = id))+
    scale_color_manual(values=c(Republican = "red", Democratic = "blue"))
  
```
```{r}
df <- tibble( #not we are just making a fake dataset so we can plot it
  x= rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() + 
  coord_fixed()
ggplot(df, aes(x,y))+
  geom_hex()+
  viridis::scale_fill_viridis()+
  coord_fixed()


```
Themes
```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_bw()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_light()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_classic()

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se = FALSE)+
  theme_dark()

theme(panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size=8),
      panel.grid.major = element_blank(),
      axis.text.y=element_text(color="black"),
      axis.text.x=element_text(color="black"),
      text=element_text(family="Arial"))

```

Saving and Exporting your Plots
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()

ggsave("my-plot.pdf")
```
Workshop 2 Assignment
```{r}
Qfish <-read.csv("C:/Users/sydne/OneDrive - James Cook University/Documents/JCU/MB5370/github/SR-Code/Module04_SR/data/QFish_Data.csv")
print(Qfish)

```
Plotting the data, It is already cleaned from my old csv
```{r}
library(ggplot2)
ggplot(Qfish, aes (x = Area, y = X2001.Total))+
  geom_bar(stat = "identity", fill = "darkcyan") +
  theme_minimal()+
  labs(title = "Total Catch in 2001 by Area",
       x = NULL,
       y = "Total Cathc") +
  theme(axis.text.x =element_text(angle = 45, hjust = 1, size = 6))


```


To Import the dataset: My first step for importing the dataset was to export a .csv file from Qfish. Once I have it in my correct file I then used the read.csv and the path to import the file into Rstudio.To make what I wanted easier to work with I created a new .csv file with only total 2001 data and the area.
Cleaning the data: it was easier to clean the data in Excel as the exported file did not make a good csv. I got my total
Creating my plot: I decided to make a bar graph as that was the best way to show the total number of species and animals caught in the year 2001. I created the boxplot and played around with color, phone sizes and axis labels. 









